---
title: "MLR Example"
author: "STAT 021 with Prof Suzy"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
 #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
#    includes:
 #     in_header: "assets/mathjax-equation-numbers.html"
    nature:
#      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')
rm(list=ls())
library('tidyverse')
library('gridExtra')
library('broom')
library('cowplot')
library("RefManageR")
library("DT")


options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE)
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```




## MLR with only numerical predictor variables 
### Public school SAT data

.scroll-output[
```{r}
SAT_data <- read_table2("Data/sat_data.txt", col_names=FALSE, cols(col_character(), col_double(), col_double(), col_double(), col_double(), col_double(), col_double(), col_double()))
colnames(SAT_data) = c("State", "PerPupilSpending", "StuTeachRatio", "Salary", "PropnStu", "SAT_verbal", "SAT_math", "SAT_tot")
SAT_data <- SAT_data %>% mutate(prop_taking_SAT = PropnStu/100) %>% select(-PropnStu)

head(SAT_data)  ##note: please dont print out entire data sets in your HW solutions!
names(SAT_data)
```
]

---
## Public school SAT data
### What are the factors affecting SAT scores? 

```{r echo=FALSE, class15_SLR, fig.align='center', out.height=400}
ggplot(SAT_data) +
  geom_point(aes(x=PerPupilSpending, y=SAT_tot)) + 
  labs(title="Public school SAT scores", subtitle = "Predictor: Per Pupil Spending",
       x="Per Pupil Spending (in thousands of dollars)", y="Total SAT Score") +
  xlim(2.5,10) + ylim(800, 1200)
```

---
## Public school SAT data
### First: Visualize 2D plots for every pair of predictors 

.scroll-output[
```{r class15_scatter, fig.align='center'}
SAT_data %>% select(-State) %>% pairs(pch=16) 
```
]


---
## Public school SAT data
### First: Visualize 2D plots for every pair of predictors 

.scroll-output[
```{r class15_scatter2, fig.align='center'}
SAT_data %>% select(-c(State, SAT_verbal, SAT_math)) %>% pairs(pch=16) 
```
]

---
## Public school SAT data
### Second: Fit a linear regression model and plot the residuals

.scroll-output[
```{r}
MLR_SAT <- lm(SAT_tot ~ PerPupilSpending + StuTeachRatio, data = SAT_data)
summary(MLR_SAT)
```
]

---
## Public school SAT data
### Second: Fit a linear regression model and plot the residuals

.scroll-output[
```{r class15_3dplot2, echo=FALSE, fig.align='center', out.height=400}
library("scatterplot3d")
require("scatterplot3d")
plot3d <- scatterplot3d(SAT_data$PerPupilSpending, 
                        SAT_data$StuTeachRatio, 
                        SAT_data$SAT_tot,
                        angle=55, scale.y=0.7, pch=16, color ="red",
                        main ="3D Regression Plane",
                        xlab="Per Pupil Spending (in thousands)", 
                        ylab="Student-teacher Ratio", 
                        zlab="Total SAT Score")
plot3d$plane3d(MLR_SAT, lty.box = "solid")
#plot3d$points3d(residual_plot_data2$PerPupilSpending, 
#                residual_plot_data2$StuTeachRatio, 
#                residual_plot_data2$fitted_vals, col="blue", type="h", pch=16)
```
]

---
## Public school SAT data
### Second: Fit a linear regression model and plot the residuals

.scroll-output[
```{r class15_resid, echo=FALSE, fig.align='center', out.height=400}
SAT_resid_data <- SAT_data %>% 
                  mutate(residuals = MLR_SAT$residuals,
                         fitted_vals = MLR_SAT$fitted.values) 
ggplot(SAT_resid_data) +
  geom_point(aes(x=fitted_vals, y=residuals)) + 
  labs(title="Residual plot", subtitle="Public school SAT scores",
       x="Predicted SAT Scores", y="Residuals") + 
  geom_hline(yintercept=0) + ylim(-200,200)
```
]

---
## Public school SAT data
### Second: Fit a linear regression model and plot the residuals


```{r eval=FALSE}
SAT_resid_data <- SAT_data %>% 
                  mutate(residuals = MLR_SAT$residuals,
                         fitted_vals = MLR_SAT$fitted.values) 
ggplot(SAT_resid_data) +
  geom_point(aes(x=fitted_vals, y=residuals)) + 
  labs(title="Residual plot", subtitle="Public school SAT scores",
       x="Predicted SAT Scores", y="Residuals") + 
  geom_hline(yintercept=0) + ylim(-200,200)
```




```{r echo=FALSE, eval=FALSE}
MLR_SAT_trans <- lm(SAT_tot ~ log(PerPupilSpending) + StuTeachRatio, data = SAT_data)
SAT_resid_data_trans <- SAT_data %>% 
                  mutate(residuals = MLR_SAT_trans$residuals,
                         fitted_vals = MLR_SAT_trans$fitted.values) 
ggplot(SAT_resid_data_trans) +
  geom_point(aes(x=fitted_vals, y=residuals)) + 
  labs(title="Residual plot", subtitle="Public school SAT scores",
       x="Predicted SAT Scores", y="Residuals") + 
  geom_hline(yintercept=0) + ylim(-200,200)
```



```{r class15_resid3, echo=FALSE, fig.align='center', eval=FALSE}
MLR_SAT_trans <- lm(SAT_tot ~ log(PerPupilSpending) + StuTeachRatio, data = SAT_data)
SAT_resid_data_trans <- SAT_data %>% 
                  mutate(residuals = MLR_SAT_trans$residuals,
                         fitted_vals = MLR_SAT_trans$fitted.values) 
trans_plot <- ggplot(SAT_resid_data_trans) +
  geom_point(aes(x=fitted_vals, y=residuals)) + 
  labs(title="Residual plot for transformed per pupil spending",
       x="Predicted SAT Scores", y="Residuals") + 
  geom_hline(yintercept=0) + ylim(-200,200)
reg_plot <- ggplot(SAT_resid_data) +
  geom_point(aes(x=fitted_vals, y=residuals)) + 
  labs(title="Residual plot for untransformed data",
       x="Predicted SAT Scores", y="Residuals") + 
  geom_hline(yintercept=0) + ylim(-200,200)

grid.arrange(grobs=list(reg_plot, trans_plot), ncol=2)
```

