---
title: "Factors in R"
author: "STAT 021 with Prof Suzy"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
 #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
#    includes:
 #     in_header: "assets/mathjax-equation-numbers.html"
    nature:
#      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')
rm(list=ls())
library('tidyverse')
library('gridExtra')
library('broom')
library('cowplot')
library("RefManageR")
library("DT")



options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE)
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 30%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```



## Categorical variables in R
### Factors

In R, the default variable type is numerical. This means that when we want to define a categorical variable we need to double check that R is interpreting it correctly. R calls categorical variables **factors**. 

If you define a vector of categorical data, make sure you put all of the different levels inside quotation marks. This makes sure that R knows the elements are not numeric. You can then use the `factor()` function in R to create a categorical variable with different levels from this non-numeric vector. 

.scroll-small[
```{r class5part2.1, echo=TRUE}
color_vector <- c("Red", "Green", "Red", "Yellow", "Green")
color_vector
color_factor <- factor(color_vector)
color_factor
```
]

---
## Categorical variables in R
### Factors

.scroll-output[
**Troubleshooting tip:** Use the `class()` function to double check a variable type. 

```{r class5part2.2, echo=TRUE}
class(color_vector)
class(color_factor) 
```


The **levels** of the categorical variable are an inherited trait from the `factor()` function. This makes it easy to create a table of our factored variable to determine the total counts within each level. 

```{r class5part2.3, echo=TRUE}
table(color_factor) 
```
]

---
## Categorical variables in R
### Factors

.scroll-output[
**Note:** Anything within quotation marks is read by R as a character, that is as as a non-numeric object. 

```{r}
num_vec <- c(1,2,2,6,2,3,5)
class(num_vec)
char_vec <- c("1","2","2","6","2","3","5")
class(char_vec)
cat_vec <- factor(char_vec)
class(cat_vec)
```

**Recall:** R is case-sensitive! So the values `"Red"` and `"red"` will be considered different levels in R.  
]



---
## Categorical variables in R
### Visualizing the data


.scroll-output[

### For categorical data, we visualize the data with a bar chart, rather than a histogram. 

Here we are going to treat the variable "number of passengers" as a categorical variable. 


```{r class4plot4, echo=TRUE, fig.align='center'}
car_data <- tibble(mph = c(22, 35, 20, 25, 24, 20, 34, 31, 30, 32, 33, 25, 29, 24),
                  passengers = c(1, 1, 3, 2, 5, 2, 1, 1, 2, 1, 3, 1, 1, 2)) 
ggplot(car_data, aes(x=factor(passengers))) +  ##notice we are applying the factor() function to the variable 'passengers' to ensure R reads this as a categorical variable
         geom_bar() + 
         labs(title="Bar chart") + 
         xlab("Number of passengers") + 
         ylab("Count")
```
]
