---
title: "Week 2: About this class and using RStudio"
author: "STAT 021 with Suzanne Thornton"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
 #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
#    includes:
 #     in_header: "assets/mathjax-equation-numbers.html"
    nature:
#      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')

rm(list=ls())
library('tidyverse')
library('gridExtra')
library('broom')
library('cowplot')
library("RefManageR")
library("DT")

#setwd("~/Google Drive Swat/Swat docs/Stat 21/Class13_files")
#setwd("~/Drive/Swat docs/Stat 21/Class9_files")
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE)
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 75%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```



# My goals for this class

- No fear of statistics! 

- It's never objective! 

- Stats is useful *everywhere* 

- Develop a critical eye 

--
- Some coding literacy and familiarity with fundamental statistical tests and models

---
# Expectations

What I expect of you as a student in my class is the following

- That you will be respectful of one another and me; 

- That you will arrive to class on-time and ready to participate; 

- That you will engage with one another in asking and answering questions (I expect lots of questions! More than I can answer by myself!);

- That you will complete the comprehension quizzes before every regular Thursday morning class;

- That you will watch the week's videos before every Tuesday class; 

- That you will communicate with me if something about the class structure isn't working for you or if you are struggling or falling behind; 

- That you will update your Zoom profile to include an image of your face and to display your preferred name (and pronouns if you wish). 

I **do not** expect everyone to have their video on all the time in Zoom. However, I do really appreciate it if you can keep your video on. It keeps me from feeling like I'm talking to myself! 


--
**Q:** What do you expect of me? 

---
# Sensitive topics in class

In this class we will study real data sets across a variety of topics. Some of the topics will be sensitive and will likely evoke some sort of emotional response. I try to be intentional with the examples we consider in class, especially on the tests. I write tests that I want you to learn something from and that I want you to feel curious about finding a solution for. 

Although statistics is widely acknowledged as a powerful tool especially in the applied sciences and in finance, I really want you to see the many ways in which statistics and data science can be useful in social justice related contexts as well. 

If there is a topic we cover that is particularly sensitive to you, please feel free to bring this up in the free-response section of the comprehension quiz for send me a DM on Slack. I will do my best to accommodate your needs and address any concerns you may have in class. 


---
# Introduction to R Studio  

- Uploading/downloading files to the Swarthmore server 

- Overview of: file types, different windows, installing and calling packages 

- Compiling HW files 

- Downloading an offline version (optional) 
  

---
# Resources for learning to program in R 

- R package `swirl` for tutorials (part of HW 1)
- Google is your friend! (and Stack Exchange, etc.)
- R for Data Science (free online textbook) https://r4ds.had.co.nz/


# Introduction to R packages

- Installing vs calling packages: the library analogy
- The R package `tidyverse` is an "umbrella-package"" that installs several other packages that are useful for data analysis, e.g. tidyr, dplyr, ggplot2, tibble.
- The functions and data that loads whenever you open R is called "Base R"


---
# Summary statistics in R 

- Mean vs median (when to use which?)


--
- Lower and upper quartiles (25th quantile/percentile, 75th quantile/percentile)


--
- Minimum and maximum (the range)


--
- Variance and standard deviation (more measures of spread)


--
.scroll-small[
```{r}
vector_of_data <- c(1,1,2,1,3,1,4,-1,-2)
mean(vector_of_data)
sd(vector_of_data)
var(vector_of_data)
summary(vector_of_data)
```
]


---
# Summary statistics in R 
## HW 1 Problem 3

"how do we know how to read the stem-and-leaf plot?"

"how could you calculate the median without using R?"

"whats the formula to calculate the standard deviation?"


Enter the data from this stem-and-leaf plot by running the code chunk below (called data_import): 
```{r data_import}
my_data <- c(-0.09, -0.09, -0.01, -0.02, -0.03, -0.04, 
             0.01, 0.01, 0.01, 0.01, 0.02, 0.03, 0.04, 0.04, 0.04, 
             0.05, 0.05, 0.05, 0.05, 0.06, 0.07, 0.09, 
             0.10, 0.10, 0.11, 0.11, 0.13, 1.00, 0.22, 0.25)
```



```{r}
## mean
## median 
## lower 25th quantile 
## lower 75th quantile 
## minimum 
## maximum 
## varaince 
## standard deviation 
```

