---
title: "Class 14 - Part 2"
subtitle: "Estimating the Error Variance"
author: "STAT 021 with Suzanne Thornton"
institute: "Swarthmore College"
date: "For class on Thursday, Oct 22, 2020 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
 #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
#    includes:
 #     in_header: "assets/mathjax-equation-numbers.html"
    nature:
#      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library('tidyverse')
library('gridExtra')
library('broom')
library('cowplot')

library("RefManageR")
library("DT")

hc_employer_2013 <- read_table2(url("http://www.swarthmore.edu/NatSci/sthornt1/DataFiles/health_care_data.txt")) 
  #read_table2(url("http://www.swarthmore.edu/NatSci/sthornt1/DataFiles/health_care_2013_cleaned.txt")) 
colnames(hc_employer_2013) <- c("Location", "TotalHealthSpending", "TotalUninsured", 
                                "TotalPop", "spending_capita", "prop_uninsured")

SLR_hc <- lm(prop_uninsured ~ spending_capita, hc_employer_2013)
#SLR_hc_pwr <- lm(prop_uninsured^(1/4) ~ spending_capita, hc_employer_2013)

#hc_resid_data <- hc_employer_2013 %>% 
#                        mutate(residuals = SLR_hc$residuals, 
#                               fitted_vals = SLR_hc$fitted.values,
#                               residuals_pwr = SLR_hc_pwr$residuals,
#                               fitted_vals_pwr = SLR_hc_pwr$fitted.values) 
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```


## What does the rest of the R output mean? 
### Residual standard error and it's degrees of freedom

Recall that $\sigma$ is also a model parameter that we don't know and must estimate. 


$$\hat{\sigma} = \text{Residual standard error} =  \sqrt{\frac{SSres}{n-2}} = \sqrt{MSE}$$

--
The *residual degrees of freedom* are $n-2$, where we are subtracting $2$ because *SSres* is computed after we have estimated two model parameters, $\beta_0$ and $\beta_1$. 


```{r}
df.residual(SLR_hc)
dim(hc_employer_2013)
```

---
## What does the rest of the R output mean? 
### Residual standard error and it's degrees of freedom

To extract this estimate for $\sigma$ from the model use: 
```{r}
SLR_hc_summary <- SLR_hc %>% summary
SLR_hc_summary$sigma
```



---
## What does the rest of the R output mean? 
### Residual standard error and it's degrees of freedom

To see that $\hat{\sigma} = \text{Residual standard error} =  \sqrt{\frac{SSres}{n-2}}$,
look at the following output: 

```{r}
anova(SLR_hc)
```


.footnote[<a href="https://campus.datacamp.com/courses/correlation-and-regression/model-fit?ex=3">Here</a> is a useful R tutorial that helps to explain what *residual standard error* is.] 



---
## What does the rest of the R output mean? 
### Residual standard error and it's degrees of freedom

.scroll-output[
```{r class14-2, fig.align='center'}
hc_resid_data <- hc_employer_2013 %>% 
                        mutate(residuals = SLR_hc$residuals, 
                               fitted_vals = SLR_hc$fitted.values) 

ggplot(hc_resid_data) +
         geom_point(aes(x=fitted_vals, y=residuals)) + 
         labs(title="Residual plot", subtitle="Cost of health care a predictor of proportion of people uninsured",
         x="Fitted values", y="Residuals") + 
  geom_hline(yintercept=0)
```


A **really good** summary of what the R output of a linear model means can be found <a href="https://stats.stackexchange.com/a/5138/77371 ">here</a>. 

**Q:** If there is evidence of heteroscedasticity, how will that affect the quality of our estimate, $\hat{\sigma}^2$? 
]
