---
title: "Week 2: Normal RVs"
author: "STAT 021 with Suzanne Thornton"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
 #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
#    includes:
 #     in_header: "assets/mathjax-equation-numbers.html"
    nature:
#      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')

rm(list=ls())
library('tidyverse')
library('ggplot2')
library('gridExtra')
library('broom')
library('cowplot')
library("RefManageR")
library("DT")


#setwd("~/Google Drive Swat/Swat docs/Stat 21/Class13_files")
#setwd("~/Drive/Swat docs/Stat 21/Class9_files")
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.path='Figs/',echo=TRUE, warning=FALSE, message=FALSE)
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 75%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```


# Normal Random Variables: Pictures and Code

Here we are going to go over how to find these attributions for a Normally distributed random variable (RV) using R code: 

- density curve/histograms

- critical value

- quantiles

- probabilities (area under the curves)

- standardization/Z-score



---
# Normal Random Variables: Pictures and Code
## Random sample

To generate a random sample from a Normal distribution with mean $3$ and standard deviation $1.5$, we use the   `rnorm` function. Below, I've labeled the sample size of $10$ as `sample_size` and then draw a random sample of that size with the specified parameters: 

```{r}
sample_size <- 10
rnorm(sample_size, mean=3, sd=1.5)
```

This is a *computer simulated* random sample. 

---
# Normal Random Variables: Pictures and Code
## Probabilities 

Given a quantile, or observation of a random variable, we can calculate how frequently this value or a smaller value will occur from this Normal distribution  with mean $3$ and standard deviation $1.5$ using the `prnorm` function. The prefix "p" stands for "probability" since this is a lower tailed probability. The example below calculates the probability a Normally distributed RV will be less than or equal to the number `my_quantile`: 

```{r}
my_quantile <- 3.45
pnorm(my_quantile, mean=3, sd=1.5, lower.tail=TRUE)
```



---
# Normal Random Variables: Pictures and Code
## Quantiles 

The *inverse* of the setting on the previous slide is to find a quantile given some probability. Given a lower tailed probability, we can calculate exactly what is the largest number for a Normally distributed RV with mean $3$ and standard deviation $1.5$ that we expect to occur this often using the `qrnorm` function. The prefix "q" stands for "quantile" since. The example below calculates the quantile or upper bound of the values of a Normal RV that occur with probability `my_probability` or less: 


```{r}
my_probability <- 0.75
qnorm(my_probability, mean=3, sd=1.5, lower.tail=TRUE)
```



---
## Normal/Gaussian Random Variables

These are continuous, quantitative random variables whose sample space is always $S= (-\infty, \infty)$.

$$X \sim N(\mu, \sigma^2)$$
The mean $(\mu)$ describes the balancing point and center of the distribution. The standard deviation $(\sigma)$ and variance $(\sigma^2)$ measure the width or spread of the histogram.

.center[<img src = "normal_hist-1.png"/>]


```{r normal_hist, echo=FALSE, fig.align='center', fig.height=5, eval=FALSE}
set.seed(101)
dat <- data.frame(x=rnorm(500))
ggplot(dat, aes(x=x)) + 
    geom_histogram(aes(y=..density..),    
                   bins=20,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  +
    scale_x_continuous(name="Quantitative variable", limits=c(-4, 4), breaks=c(-3,-2,-1,0,1,2,3), labels = c("\u03bc - 3 \u03c3","\u03bc-2\u03c3","\u03bc-\u03c3","\u03bc","\u03bc+\u03c3","\u03bc+2\u03c3","\u03bc+3\u03c3")) +
    scale_y_continuous(name="Density/Relative Frequency", limits=c(0, 0.5))

```


---
## Normal/Gaussian Random Variables

$$X \sim N(\mu, \sigma^2)$$

.center[<img src = "normal_hist2-1.png"/>]

```{r normal_hist2, echo=FALSE, fig.align='center', fig.height=5, eval=FALSE}
set.seed(101)
dat <- data.frame(x=rnorm(500))
ggplot(dat, aes(x=x)) + 
    geom_histogram(aes(y=..density..),    
                   bins=20,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  +
    scale_x_continuous(name="Quantitative variable", limits=c(-4, 4), breaks=c(-3,-2,-1,0,1,2,3), labels = c("\u03bc - 3 \u03c3","\u03bc-2\u03c3","\u03bc-\u03c3","\u03bc","\u03bc+\u03c3","\u03bc+2\u03c3","\u03bc+3\u03c3")) +
    scale_y_continuous(name="Density/Relative Frequency", limits=c(0, 0.5))

```

For Normally distributed RVs we have the following rule: $68\%$ of all possible values this RV can take on fall within **one** standard deviation of the mean, $95\%$  of all possible values fall within **two** standard deviations of the mean, and $99.7\%$ of all possible values fall within **three** standard deviation of the mean 



---
## Normal/Gaussian Random Variables
### Standard Normal RV

$$Z \sim N(0,1)$$ 


### In R some useful functions for Normal RVs are
```{r, eval=FALSE}
rnorm(100, mean=2, sd=0.9)  ## This generates 100 independence random samples from a Normal distribution with mean 2 and variance 0.9^2
pnorm(1.1, mean=2, sd=0.9, lower.tail=FALSE)  ## This find the probability that a Normally distributed (mean 2, variance 0.9^2) RV take on a value of 1.1 or higher  
qnorm(0.18, mean=2, sd=0.9, lower.tail=TRUE)  ## this finds the lower 18th percentile for a RV with a Normal distribution (mean 2, variance 0.9^2)
```


---
## Properties of Normal RVs


**1) Linear combinations of Normal RVs**
If $X \sim N(\mu_{X}, \sigma^2_{X})$ and $Y \sim N(\mu_{Y}, \sigma^2_{Y})$, then for any constant numbers $a,b,$ and $c$, 
$$aX + bY + c \sim N(\mu_{combined}, \sigma^2_{combinded}), \quad\text{where}$$
$$\mu_{combined} = a\mu_{X} + b\mu_{Y}+c \quad \text{and} \quad \sigma^2_{combined} = \sigma^2_{X} + \sigma^2_{Y} + Cov(X,Y).$$



**2) Unique specifications of Normal RVs**

The mean $(\mu)$ and variance $(\sigma^2)$ of a Normal RV completely and uniquely specify a Normal RV. This means if two random variables are both Normally distributed and have the same mean and same variance, then they are from the exact same distribution. 
