---
title: "Stat 021 Homework 5"
author: "Kayonna L Tindle"
date: 'Due: Friday, Nov. 1, 12:00pm'
output:
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(magrittr)
```


```{css, echo=FALSE}
.red{color: #ce151e;}
```

<center>**Instructions:** A **pdf** version of your homework must be submitted to Gradescope by **noon** on the due date. The course passcode is **MPKJ4Z**. If you are having trouble getting your *.Rmd* file to compile, you need to get help with this  **before** the due date. 

You are allowed to hand in **only one** late homework assignment throughout the semester. If you need to hand in this particular assignment late, you must let me know via email by noon on the due date.

You are encouraged to study with your peers to help complete the homework assignments but no copying is allowed. If I see that two or more homework assignments are copied, all students involved will receive a grade of $0$ on that assignment and will forfeit (perhaps retroactively) the opportunity to hand in a late homework. </center> 

***

**Q1)** Sketch (by hand) residual plots (with $\hat{y}_i$, predicted response values, on the horizantal axis) that show each of the following: (5 points)
  1. constant variance and linearity; 
  1. non-constant variance and linearity; 
  1. constant variance and non-linearity;
  1. non-constant variance and non-linearity. 
  
  
**Q2)** Suppose we have two random variables $X$ and $Y$. What are the differences among the following assumptions regarding $X$ and $Y$: 

- $X$ and $Y$ are uncorrelated, 
  
- $X$ and $Y$ are independent, 
  
- $X$ and $Y$ have the same variance, and 
  
- $X$ and $Y$ have the same distribution? (5 points)


**Q3)** Read the Wikipedia page for Simpson's Paradox: https://en.wikipedia.org/wiki/Simpson%27s_paradox. Then, import the "Stand your ground" data set uploaded on Moodle. This data (from 2015) is related to the Stand Your Ground law in Florida. Each observational unit consists of a case where the Stand Your Ground law was a part of the defense strategy, the defendent's race (white or non-white), the victim's race (white or non-white), and a binary variable indicating whether or not the defendent was convicted. With this categorical data we are not going to fit a regression model but we are going to examine this data and look out for Simpson's paradox. (10 points)

a) Create and print the following tables to summarize the data: 

    1. Defendant's race vs convicted for all observational units; 
    
    1. Defendant's race vs convicted for cases with minority victims only; 

    1. Defendant's race vs convicted for cases with white victims only; 

    1. The table created by adding Tables 2 and 3 together.
  
b) What are the overall conviction rates for minority and white defendants, respectively? What are the conviction rates for minority and white defendants among the cases with minority victims? What are the conviction rates for minority and white defendants among the cases with white victims? 

c) Explain what is going on here in terms of Simpson's paradox and interpret what this means with respect to racial bias in the criminal justice system. 

**Q2**
Suppose we have two random variables $X$ and $Y$. What are the differences among the following assumptions regarding $X$ and $Y$: 

- $X$ and $Y$ are uncorrelated, this idea relates back to covariance wherein if a two varables are uncorrelated, we cannot fit a linear relationship to the two variables. We no that there the covariance between the two random variables equals zero. 
  
- $X$ and $Y$ are independent, this idea relates back more strongly to probabilty wherein we could state that the probability of $X$ and $Y$ combined is the same (in other words, does "not" change). More plainly, the outcome of one variable, $X$, does not affect the results of variable $Y$. When two variables are independent, they will be inherently uncorrelated as well.
  
- $X$ and $Y$ have the same variance, and $X$ and $Y$ have the same distribution?
Observing equal variance relates to not having heteroscadicity, wherein the spread of the data points are about equal across the data set. However, in relation to random variables $X$ and $Y$, we will observe similar spread between the data sets wherein variance of both random variables may equal 1. Nonetheless, when we observe distribution in the data set, we are displaying the values of a data set and how frequently we will observe those values. Since variance is not the only factor that goes into a distribution and its shape (i.e we also look at the mean). we can not say that two random variables with the same variance will have the same distribution. (5 points)


```{r Q3}
setwd("~/Stat 21 HW")
read_table2("stand_you_ground.txt")
dtable1 <- read_table2("stand_you_ground.txt")
print(dtable1)
View(dtable1)
dtable2 <- dtable1
dtable2 <- dtable1 %>% select(Convicted, Accused, MinVictim) %>% filter(MinVictim==1)
print(dtable2)
View(dtable2)
dtable3 <- dtable1
dtable3 <- dtable1 %>% select(Convicted, Accused, WhiteVictim) %>% filter(WhiteVictim==1)
print(dtable3)
View(dtable3)

# Conviction rate defined: "conviction rate of a prosecutor or government is the number of convictions divided by the number of criminal cases brought." -Wikipedia
# overall conviction rates for minority and white defendants, respectively
# percent of cases brought of white defendants:  divided by convictions is 34 percent
q3b1 <- dtable1 %>% filter(Accused=="White", Convicted=="Yes")
q3b2 <- dtable1 %>% filter(Accused=="White")
count(q3b1)/count(q3b2)
# percent of cases brought of minority defendants:  divided by convictions is 33 percent 
q3b3 <- dtable1 %>% filter(Accused=="Minority", Convicted=="Yes")
q3b4 <- dtable1 %>% filter(Accused=="Minority")
count(q3b3)/count(q3b4)
# percent the conviction rates for minority and white defendants among the cases with minority victims is 27 percent
q3b5 <- dtable1 %>% filter(MinVictim=="1", Convicted=="Yes")
q3b6 <- dtable1 %>% filter(MinVictim=="1")
count(q3b5)/count(q3b6)
# percent the conviction rates for minority and white defendants among the cases with white victims is 38 percent
q3b7 <- dtable1 %>% filter(WhiteVictim=="1", Convicted=="Yes")
q3b8 <- dtable1 %>% filter(WhiteVictim=="1")
count(q3b7)/count(q3b8)
# part c. these findings apply to the Simpson's paradox by showing how if we were to observe the data set of conviction rates when our data is organized in a way that has our data group of victims "WhiteVictim" and "MinVictim" combined, we would see a generalized trend that does not reflect the noteable difference (of apprximately 10 percent) in how conviction rates for cases where the stand your ground law is applied when the victim in the case is white compared to when the victim in the case is minority. If we actually evaluate the data by distinguishing or grouping the victims variable by white or minority, we will see cause to consider bias in how cases are handled on this law when the victim is white and not minority (when minority victim, the case is less likely to result in a conviction). 
```